On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-09-07 paper.tex
Wording

\begin{theorem}
\label{theo: funt, form}
For $a, b \in \mathbb R^+$, $a \geq b$ and $\len \in \mathbb R^+$,
\begin{equation}
\label{eq: theo: funt, form; funt}
\funt(\len) = \isoli+\jsoli-1
\end{equation}
with
\begin{align}
\label{eq: theo: funt, form; i}
\isoli &= \left\lceil \frac{a + b \Re \sqrt{4\len^2/(a^2+b^2)-1}}{2a} \right\rceil + 1, \\
\label{eq: theo: funt, form; j}
\jsoli &= \left\lceil \frac{\sqrt{\len^2-(\isoli-2)^2a^2}}{b} \right\rceil + 1.
\end{align}
The function $\funt$ is piecewise constant and left-continuous, with unit-height jumps. A jump occurs at $\len$ if and only if $\len = \funl(\tiles)$ for some $\tiles \in \mathbb N, \tiles \geq 4$; and then $\funt(\len) = \tiles-1$, $\lim_{h \rightarrow 0+} \funt(\len+h) = \tiles$.
\end{theorem}

\begin{proof}
The idea of the proof is loosely similar to that used in Proposition~\ref{prop: min suff set, form}:
% ***is it appropriate to say that it is loosely similar? Both are based on a continuous approximation which is then quantized. Does ``loosely'' have a negative connotation?
find the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ for $i, j \geq 2$, and from that obtain the actual pair $(\isoli, \jsoli)$ of integer numbers that yields the solution.

Specifically, assuming first that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. This value, $\isoli$, is computed by rounding down $\isolr$. Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained. As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in $\mss$, and in other cases it is not. However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$. Thus the desired result is $\isoli+\jsoli-1$.

@@

As for the properties of $\funt$, it follows from \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j} that it is piecewise constant and left-continuous. From the procedure described in the previous paragraphs for obtaining $(\isoli, \jsoli)$ it is clear that $\isoli+\jsoli-1$ increases in steps of $1$ when $\len$ is increased continuously; that is, $\funt$ has jumps of unit value.

Consider an arbitrary $\len$ such that for some $\tiles \in \mathbb N, \tiles \geq 4$
\begin{equation}
\label{eq: theo funt, form: prop 1}
\funl(\tiles)=\len.
\end{equation}
To see that $\funt$ has a jump at $\len$, assume for the sake of contradiction that $\funt$ is continuous, therefore constant, at that point. Then there is $\epsilon > 0$ for which $\funt(\len+\epsilon) = \tiles = \funt(\len-\epsilon)$. This means that there exists a segment with length $\len-\epsilon$ that visits $\tiles$ tiles, and thus $\funl(\tiles) \leq \len-\epsilon < \len$, in contradiction with \eqref{eq: theo funt, form: prop 1}. Therefore $\funt$ is discontinuous, from the right, at $\len$. By definition of $\funl$, \eqref{eq: theo funt, form: prop 1} implies that $\funt(\len) < \tiles$, and that there exists $\epsilon > 0$ such that $\funt(\len+h) = \tiles$ for $0 < h < \epsilon$. This means that
\begin{align}
\label{eq: theo funt, form: prop 2}
\funt(\len) &< \tiles, \\
\label{eq: theo funt, form: prop 3}
\lim_{h \rightarrow 0+} \funt(\len+h) &= \tiles,
\end{align}
that is, $\funt$ has a jump at $\len$. In addition, since the jump has unit height, it follows from \eqref{eq: theo funt, form: prop 2} and \eqref{eq: theo funt, form: prop 3}  that $\funt(\len) = \tiles-1$.

@@

By Theorem~\ref{theo: funt, form}, $\funt$ has unit-height jumps at the values $\funl(\tiles)$, $\tiles \in \mathbb N, \tiles \geq 4$. This implies that $\funl$ is monotone increasing for $\tiles \geq 3$.
\end{proof}

>>>>

\begin{theorem}
\label{theo: funt, form}
For $a, b \in \mathbb R^+$, $a \geq b$ and $\len \in \mathbb R^+$,
\begin{equation}
\label{eq: theo: funt, form; funt}
\funt(\len) = \isoli+\jsoli-1
\end{equation}
with
\begin{align}
\label{eq: theo: funt, form; i}
\isoli &= \left\lceil \frac{a + b \Re \sqrt{4\len^2/(a^2+b^2)-1}}{2a} \right\rceil + 1, \\
\label{eq: theo: funt, form; j}
\jsoli &= \left\lceil \frac{\sqrt{\len^2-(\isoli-2)^2a^2}}{b} \right\rceil + 1.
\end{align}
The function $\funt$ is piecewise constant and left-continuous, with unit-height jumps. A jump occurs at $\len$ if and only if $\len = \funl(\tiles)$ for some $\tiles \in \mathbb N$, $\tiles \geq 4$; and then $\funt(\len) = \tiles-1$, $\lim_{h \rightarrow 0+} \funt(\len+h) = \tiles$.
\end{theorem}

\begin{proof}
The approach is similar to that used in Proposition~\ref{prop: min suff set, form}. Namely, for $i, j \geq 2$ the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ is found, and from that the actual integer solution $(\isoli, \jsoli)$ is computed.

More specifically, assuming that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. This value, $\isoli$, is computed by rounding down $\isolr$. Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained. As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in $\mss$, and in other cases it is not. However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$. Thus the desired result is $\isoli+\jsoli-1$.

@@

As for the properties of $\funt$, it follows from \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j} that this function is piecewise constant and left-continuous. From the procedure described in the previous paragraphs for obtaining $(\isoli, \jsoli)$ it is clear that $\isoli+\jsoli-1$ increases in steps of $1$ when $\len$ is increased continuously; that is, $\funt$ has jumps of unit height.

Consider an arbitrary $\len$ such that for some $\tiles \in \mathbb N$, $\tiles \geq 4$
\begin{equation}
\label{eq: theo funt, form: prop 1}
\funl(\tiles)=\len.
\end{equation}
To see that $\funt$ has a jump at $\len$, assume for the sake of contradiction that $\funt$ is continuous, therefore constant, at that point. Then there is $\epsilon > 0$ for which $\funt(\len+\epsilon) = \tiles = \funt(\len-\epsilon)$. This means that there exists a segment with length $\len-\epsilon$ that visits $\tiles$ tiles, and thus $\funl(\tiles) \leq \len-\epsilon < \len$, in contradiction with \eqref{eq: theo funt, form: prop 1}. Therefore $\funt$ is discontinuous, from the right, at $\len$. By definition of $\funl$, \eqref{eq: theo funt, form: prop 1} implies that $\funt(\len) < \tiles$, and that there exists $\epsilon > 0$ such that $\funt(\len+h) = \tiles$ for $0 < h < \epsilon$. This means that
\begin{align}
\label{eq: theo funt, form: prop 2}
\funt(\len) &< \tiles, \\
\label{eq: theo funt, form: prop 3}
\lim_{h \rightarrow 0+} \funt(\len+h) &= \tiles,
\end{align}
that is, $\funt$ has a jump at $\len$. In addition, since the jump has unit height, it follows from \eqref{eq: theo funt, form: prop 2} and \eqref{eq: theo funt, form: prop 3}  that $\funt(\len) = \tiles-1$.

@@

By Theorem~\ref{theo: funt, form}, $\funt$ has unit-height jumps at the values $\funl(\tiles)$, $\tiles \in \mathbb N$, $\tiles \geq 4$. This implies that $\funl$ is monotone increasing for $\tiles \geq 3$.
\end{proof}
